---
title: "01_Exploratory"
author: "Anjali Gupta"
date: "4/5/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading Dyes Data


```{r}
mydata <- read.csv('dyes.csv', header=TRUE)

table(mydata$color)
table(mydata$color, mydata$sample)
(unique(mydata$sample)) # 20 unique samples
(unique(paste(mydata$color, mydata$sample))) # 30 unique combinations of color and sample, 10 for each color

#Add another variable in the dataset
#Unique combination of dye color and sample number
id <- paste(mydata$color, mydata$sample, sep = "")
name <- mydata$name
mydata <- cbind("name" = mydata[,1], id, mydata[,2:ncol(mydata)])
```

## Plots


```{r}
pdf("FDObservation1.pdf", width = 8.27, height = 5.83)
plot(1:3201, mydata[1,6:3206], type='l', xlab = "", ylab = "", main = "Single Functional Observation", col = "cyan", lwd = 2)
title(ylab="Absorbance", line=2.5, cex.lab=1)
title(xlab="Wavenumber", line=2.5, cex.lab=1)
graphics.off()

mydata_t <- t(mydata)

#Separate out C, Y, M spectra
mydataC <- mydata[mydata$color == "C",]
mydataY <- mydata[mydata$color == "Y",]
mydataM <- mydata[mydata$color == "M",]

#Transpose the data
mydataC_t <- t(mydataC)
colnames(mydataC_t) <- mydataC_t[1,]
mydataC_t <- mydataC_t[- c(1:5),]
mydataC_t <- cbind(1:nrow(mydataC_t), mydataC_t)
class(mydataC_t) <- "numeric"

mydataY_t <- t(mydataY)
colnames(mydataY_t) <- mydataY_t[1,]
mydataY_t <- mydataY_t[- c(1:5),]
mydataY_t <- cbind(1:nrow(mydataY_t), mydataY_t)
class(mydataY_t) <- "numeric"

mydataM_t <- t(mydataM)
colnames(mydataM_t) <- mydataM_t[1,]
mydataM_t <- mydataM_t[- c(1:5),]
mydataM_t <- cbind(1:nrow(mydataM_t), mydataM_t)
class(mydataM_t) <- "numeric"

#Line plots
pdf("cyan.pdf",width = 8.27, height = 5.83)
plot(mydataC_t[,1:2], type='l',col="cyan", lwd = 1, main = "Raman Spectra: Cyan Inks", xlab = "", ylab = "") #check x- and y-axis labels with James
title(ylab="Absorbance", line=2.5, cex.lab=1)
title(xlab="Wavenumber", line=2.5, cex.lab=1)
for(i in 3:ncol(mydataC_t)){
  lines(mydataC_t[,1],mydataC_t[,i], col="cyan", lwd = 1)
}
graphics.off()
#Identifying peculiar lines
pdf("cyan_details_v2.pdf",width = 8.27, height = 5.83)
plot(mydataC_t[,1:2], type='l',col="cyan", lwd = 1, main = "Raman Spectra: Cyan Inks", xlab = "", ylab = "") #check x- and y-axis labels with James
title(ylab="Absorbance", line=2.5, cex.lab=1)
title(xlab="Wavenumber", line=2.5, cex.lab=1)
for(i in 3:ncol(mydataC_t)){
  lines(mydataC_t[,1],mydataC_t[,i], col="cyan", lwd = 1)
}
#which(mydataC_t[1:10,] > 0.6, arr.ind=T) #C.440.4
lines(mydataC_t[,1],mydataC_t[,45], col="cyan4", lwd = 2, lty = 2) #C.440.4
#which(mydataC_t[1775,] > 1, arr.ind = T) #C.372.4
#lines(mydataC_t[,1],mydataC_t[,20], col="cyan3", lwd = 2, lty = 3) #C.372.4
lines(mydataC_t[,1],mydataC_t[,20], col="cadetblue4", lwd = 2, lty = 6) #C.372.4
legend("topright", col = c("cyan4", "cadetblue4", "cyan"), lwd = c(2,2,1), lty = c(2,6,1), legend = c("C.440.4", "C.372.4", "Other Cyan Inks"), bty="n")
graphics.off()

#plot(mydataC_t[,1:2], type='l',col="cyan", lwd = 2, main = "Single Functional Observation", xlab = "Relative Wavenumber", ylab = "Normalised Intensity")

pdf("FDObservationsC37_1.pdf", width = 8.27, height = 5.83)
plot(mydataC_t[,1:2], type='l',col="cyan", lwd = 1, main = "Functional Observations - Ink C37", xlab = "", ylab = "")
title(ylab="Absorbance", line=2.5, cex.lab=1)
title(xlab="Wavenumber", line=2.5, cex.lab=1)
for(i in 3:6){
  lines(mydataC_t[,1],mydataC_t[,i], col="cyan", lwd = 1)
}
graphics.off()

pdf("yellow_v2.pdf", width = 8.27, height = 5.83)
plot(mydataY_t[,1:2], type='l',col="goldenrod1", lwd = 1, main = "Raman Spectra: Yellow Inks", xlab = "", ylab = "")
title(ylab="Absorbance", line=2.5, cex.lab=1)
title(xlab="Wavenumber", line=2.5, cex.lab=1)
for(i in 3:ncol(mydataY_t)){
  lines(mydataY_t[,1],mydataY_t[,i], col="goldenrod1", lwd = 1)
}
graphics.off()

pdf("magenta_v2.pdf", width = 8.27, height = 5.83)
plot(mydataM_t[,1:2], type='l',col="magenta", lwd = 1, main = "Raman Spectra: Magenta Inks", xlab = "", ylab = "")
title(ylab="Absorbance", line=2.5, cex.lab=1)
title(xlab="Wavenumber", line=2.5, cex.lab=1)
for(i in 3:ncol(mydataM_t)){
  lines(mydataM_t[,1],mydataM_t[,i], col="magenta", lwd = 1)
}
graphics.off()
```

```{r Refund.wave}
# library(refund.wave)
# 
# data(gasoline)
# gas.wnet1 <- wnet(gasoline$octane, xfuncs = gasoline$NIR[,1:256],
# nfeatures= 20, min.scale = 0, alpha = 1)
# gas.wpcr1 <- wcr(gasoline$octane, xfuncs = gasoline$NIR[,1:256], min.scale = 0,
# nfeatures = 20, ncomp = 15)
# gas.wpls1 <- wcr(gasoline$octane, xfuncs = gasoline$NIR[,1:256], min.scale = 0,
# nfeatures = 20, ncomp = 15, method = "pls")
# plot(gas.wnet1)
# plot(gas.wpcr1)
# plot(gas.wpls1)
```

## Create FDA objects and plot

Plot functional data objects and plot them along with the observed data.

```{r fda}
library('fda')

basisfd <- create.bspline.basis(rangeval=c(1, 3201), nbasis=NULL, norder=5, breaks=seq(1, 3201, 20))

#Fitting Spline basis to the functional data
for(i in 1:length(unique(id))){
    y1 <- t(mydata[mydata$id == unique(id)[i],6:3206])

    y1.fd <- Data2fd(y1, argvals = 1:3201, basisfd, fdnames=list("Wavenumber","rep","Absorbance"))

    plot.fd(y1.fd, main=unique(id)[i], ylim = c(0,2))
    for(k in 1:ncol(y1)){
      lines(y1[,k], col="gray")
    }
}
rm(y1.fd)
rm(y1)

##Single spectrum fitted using functional objects
i <- 80
y <- t(mydata[i,6:3206])
y.fd <- Data2fd(y, argvals = 1:3201, basisfd, fdnames=list("Wavenumber","rep","Absorbance"))
pdf(paste(name[i], "_fda_v2.pdf", sep=""), width = 8.27, height = 5.83)
plot.fd(y.fd, main=name[i], ylim = c(0,2), lwd = 2, col = "red", xlab="",ylab="")
title(ylab="Absorbance", line=2.5, cex.lab=1)
title(xlab="Wavenumber", line=2.5, cex.lab=1)
lines(y, col="black", lwd = 2, lty = 3)
legend("topright", col = c("black", "red"), lwd = 2, legend = c("Original Spectrum", "Functional Data Object"), bty = "n", lty = c(3, 1))
graphics.off()

par(mfrow=c(2,2))
i <- c(1, 10, 80, 147)
for(k in 1:length(i)){
y <- t(mydata[i[k],6:3206])
y.fd <- Data2fd(y, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))
plot.fd(y.fd, main=name[i[k]], ylim = c(0,2), lwd = 2, col = "red")
lines(y, col="black", lwd = 2, lty = 3)
legend("topright", col = c("black", "red"), lwd = 2, legend = c("Original Spectrum", "Functional Data Object"), bty = "n", lty = c(3, 1))
}
dev.off()

```

## Test 1 - Permutation Test for Functional Hypothesis - Functional T test

This test computes t statistic at each point of the spectrum and the final test statistic is the maximum of the computed test statistics.

```{r Test1}
#Permutation Test of Functional Hypothesis - Functional T test
pvals <- matrix(NA, nrow=length(unique(id)), ncol=length(unique(id)))
rownames(pvals) <- (unique(id))
colnames(pvals) <- unique(id)

for(i in 1:(length(unique(id)) - 1)){#1:
  y1 <- t(mydata[mydata$id == unique(id)[i],6:3206])
  
  y1.fd <- Data2fd(y1, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))
  
  for(j in (i+1):length(unique(id))){
    y2 <- t(mydata[mydata$id == unique(id)[j],6:3206])
    
    #Test to check the p-value of T-test between replications from the same sample
    # y1.temp <- y1[,1:3]
    # y2.temp <- y1[,4:5]
    # y1.temp.fd <- Data2fd(y1.temp, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))
    # y2.temp.fd <- Data2fd(y2.temp, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))
    # tperm.fd(y1.temp.fd, y2.temp.fd)

    y2.fd <- Data2fd(y2, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))
    
    permtest <- tperm.fd(y1.fd, y2.fd)
    
    pvals[j,i] <- permtest$pval
    
   #  plot.fd(y1.fd, main=unique(id)[i], ylim = c(0,2))
   #  for(k in 1:ncol(y1)){
   #    lines(y1[,k], col="gray")
   #  }
   #  
   # plot.fd(y2.fd, main=unique(id)[j], ylim = c(0,2))
   #  for(k in 1:ncol(y2)){
   #    lines(y2[,k], col="gray")
   #  }
  }
}

write.table(pvals, "PermutationTTest.txt", quote=FALSE, sep="\t", col.names = TRUE, row.names = TRUE, dec = ".")

```

## Test 2 - Compare the coefficients of the basis functions individually

```{r Test2}
# Comparing the coefficients of the basis functions - individually
library('fdatest')
data1 <- mydata[mydata$id == unique(id)[1], 6:3206]
data2 <- mydata[mydata$id == unique(id)[2], 6:3206]
test12 <- ITP2bspline(data1, data2, order = 5, nknots = length(seq(1, 3201, 20)))
# Plotting the results of the ITP
plot(test12) #,main='NASA data',xrange=c(1,365),xlab='Day'
# Plotting the p-values heatmap
ITPimage(test12)#,abscissa.range=c(0,12)
# Selecting the significant components at 5% level
which(test12$corrected.pval < 0.05)

# Compare the results with fda - tperm.fd - Compares the objects at argvals = 101 by default i.e. 101 values from the whole spectra.
# basisfd <- create.bspline.basis(rangeval=c(1, 3201), nbasis=NULL, norder=5, breaks=seq(1, 3201, 20))
# data1_t <- t(data1)
# data2_t <- t(data2)
# data1.fd <- Data2fd(data1_t, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))
# data2.fd <- Data2fd(data2_t, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))
# test12_perm <- tperm.fd(data1.fd, data2.fd)

```

## Test 3 Trial - Hotelling T2

```{r Test3}
# FDA - Hotelling T2
#install.packages("devtools")
#library(devtools)
#install_github("astamm/fdahotelling")
library(fdahotelling)

basisfd <- create.bspline.basis(rangeval=c(1, 3201), nbasis=NULL, norder=5, breaks=seq(1, 3201, 20))
zero <- matrix(0, nrow = 5, ncol = 3201)
zero_t <- t(zero)
zero.fd <- Data2fd(zero_t, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))

#test12_hot <- test_twosample(x = data1, y = data2, step_size = 0.01, mu = 0, B = 0)

test12_hot.fd <- test_twosample(x = data1.fd, y = data2.fd, step_size = 0.01, mu = zero.fd, B = 0)

# Running for all combinations
pvals <- matrix(NA, nrow=length(unique(id)), ncol=length(unique(id)))
rownames(pvals) <- (unique(id))
colnames(pvals) <- unique(id)

for(i in 1:(length(unique(id)) - 1)){#1:
  y1 <- t(mydata[mydata$id == unique(id)[i],6:3206])
  
  y1.fd <- Data2fd(y1, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))
  
  for(j in (i+1):length(unique(id))){
    y2 <- t(mydata[mydata$id == unique(id)[j],6:3206])
    
    #Test to check the p-value of T-test between replications from the same sample
    # y1.temp <- y1[,1:3]
    # y2.temp <- y1[,4:5]
  #  y1.temp <- y1[,1:2]
  #  y2.temp <- y1[,3:4]
    # y1.temp.fd <- Data2fd(y1.temp, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))
    # y2.temp.fd <- Data2fd(y2.temp, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))
    # test_hot.fd <- test_twosample(x = y1.temp.fd, y = y2.temp.fd, step_size = 0.01, mu = zero.fd, B = 100)
    # pvals[i,i] <- test_hot.fd$pValue
    
    y2.fd <- Data2fd(y2, argvals = 1:3201, basisfd, fdnames=list("Wave Number","rep","Normalized Intensity"))
    
    #test_hot.fd <- test_twosample(x = y1.fd, y = y2.fd, step_size = 0.01, mu = zero.fd, B = 0) #Parametric Test
    test_hot.fd <- test_twosample(x = y1.fd, y = y2.fd, step_size = 0.01, mu = zero.fd, B = 1000) #Permutation Test
    
    pvals[j,i] <- test_hot.fd$pValue
    

  }
}

write.table(pvals, "HotellingT2Test-NonParametric.txt", quote=FALSE, sep="\t", col.names = TRUE, row.names = TRUE, dec = ".")
```

## Example from 'fdahotelling' package - fd objects

```{R}
# An fd object is created from the transposed matrix 
# in the sense that it expects observations to be stored
# in columns and grid points in rows.
lower_ind <- which(aneurisk$variable == "radius" & aneurisk$group == "low")
upper_ind <- which(aneurisk$variable == "radius" & aneurisk$group == "up")
arc_length_lower <- aneurisk$abscissa[[lower_ind]]
arc_length_upper <- aneurisk$abscissa[[upper_ind]]
# Both datasets must be defined on the same grid
all(arc_length_upper == arc_length_lower)
#> [1] TRUE
arc_length <- arc_length_upper
# Effective transformation to fd objects
data_lower <- aneurisk$data[[lower_ind]]
fd_lower <- fda::Data2fd(arc_length, t(data_lower))
data_upper <- aneurisk$data[[upper_ind]]
fd_upper <- fda::Data2fd(arc_length, t(data_upper))
# Requirement: the differences between the mean assumed 
# in the null hypothesis must be converted in fd format 
# as well.
fd_delta <- fda::Data2fd(arc_length, matrix(0, length(arc_length), dim(data_upper)[1]))

set.seed(1234)
test_twosample(
  x = aneurisk$data[[lower_ind]], 
  y = aneurisk$data[[upper_ind]], 
  step_size = 0.01, 
  B = 100L
)
#  - P-value resolution: 0.01
#  - Computing approximate p-value using 100 random permutations.
#  - P-value will not drop below 7.91072860244862e-15 in average.
# # A tibble: 1 x 3
#   statName  statVal pValue
#   <chr>       <dbl>  <dbl>
# 1 Hotelling    385.   0.96

set.seed(1234)
test_twosample(
  x = fd_lower, 
  y = fd_upper, 
  mu = fd_delta,
  step_size = 0.01, 
  B = 100L
)
#  - P-value resolution: 0.01
#  - Computing approximate p-value using 100 random permutations.
#  - P-value will not drop below 7.91072860244862e-15 in average.
# # A tibble: 1 x 3
#   statName  statVal pValue
#   <chr>       <dbl>  <dbl>
# 1 Hotelling    526.   0.71
  
set.seed(1234)
test_twosample(
  x = t(fd_lower$coefs), 
  y = t(fd_upper$coefs), 
  step_size = 0.01, 
  B = 100L
)
#  - P-value resolution: 0.01
#  - Computing approximate p-value using 100 random permutations.
#  - P-value will not drop below 7.91072860244862e-15 in average.
# # A tibble: 1 x 3
#   statName  statVal pValue
#   <chr>       <dbl>  <dbl>
# 1 Hotelling    404.   0.94

## Derivates --
set.seed(1234)
lower_ind <- which(aneurisk$variable == "radius_der" & aneurisk$group == "low")
upper_ind <- which(aneurisk$variable == "radius_der" & aneurisk$group == "up")
test_twosample(
  x = aneurisk$data[[lower_ind]], 
  y = aneurisk$data[[upper_ind]], 
  step_size = 0.01, 
  B = 100L
)
#  - P-value resolution: 0.01
#  - Computing approximate p-value using 100 random permutations.
#  - P-value will not drop below 7.91072860244862e-15 in average.
# # A tibble: 1 x 3
#   statName  statVal pValue
#   <chr>       <dbl>  <dbl>
# 1 Hotelling   5350.   0.02
#  - P-value resolution: 0.001
#  - Computing approximate p-value using 1000 random permutations.
#  - P-value will not drop below 7.91072860244862e-15 in average.
# # A tibble: 1 x 3
#   statName  statVal pValue
#   <chr>       <dbl>  <dbl>
# 1 Hotelling   5350.  0.011
arc_length_lower <- aneurisk$abscissa[[lower_ind]]
arc_length_upper <- aneurisk$abscissa[[upper_ind]]
# Both datasets must be defined on the same grid
all(arc_length_upper == arc_length_lower)
#> [1] TRUE
arc_length <- arc_length_upper
# Effective transformation to fd objects
data_lower <- aneurisk$data[[lower_ind]]
fd_lower <- fda::Data2fd(arc_length, t(data_lower))
data_upper <- aneurisk$data[[upper_ind]]
fd_upper <- fda::Data2fd(arc_length, t(data_upper))
set.seed(1234)
test_twosample(
  x = fd_lower, 
  y = fd_upper, 
  mu = fd_delta,
  step_size = 0.01, 
  B = 100L
)
#  - P-value resolution: 0.01
#  - Computing approximate p-value using 100 random permutations.
#  - P-value will not drop below 7.91072860244862e-15 in average.
# # A tibble: 1 x 3
#   statName  statVal pValue
#   <chr>       <dbl>  <dbl>
# 1 Hotelling   7097.   0.02
#  - P-value resolution: 0.001
#  - Computing approximate p-value using 1000 random permutations.
#  - P-value will not drop below 7.91072860244862e-15 in average.
# # A tibble: 1 x 3
#   statName  statVal pValue
#   <chr>       <dbl>  <dbl>
# 1 Hotelling   7097.  0.015
set.seed(1234)
test_twosample(
  x = t(fd_lower$coefs), 
  y = t(fd_upper$coefs), 
  step_size = 0.01, 
  B = 100L
)
#  - P-value resolution: 0.01
#  - Computing approximate p-value using 100 random permutations.
#  - P-value will not drop below 7.91072860244862e-15 in average.
# # A tibble: 1 x 3
#   statName  statVal pValue
#   <chr>       <dbl>  <dbl>
# 1 Hotelling   4995.   0.02
#  - P-value resolution: 0.001
#  - Computing approximate p-value using 1000 random permutations.
#  - P-value will not drop below 7.91072860244862e-15 in average.
# # A tibble: 1 x 3
#   statName  statVal pValue
#   <chr>       <dbl>  <dbl>
# 1 Hotelling   4995.   0.01
```